---
title: "endemism"
format: html
editor: visual
---

## Endemism

We love endemism

```{r}
#install packages


#load libraries
library(tidyr)
library(tidyverse)
library(data.table)
library(dplyr)
library(ggplot2)

#read csvs
genus_end <- read.csv("data/australian_flora/genus_level_endemism_estimate.csv")

#wcvp <- fread(file = "data/wcvp/wcvp_names.csv", header = T)
#wcvp_distribution <- fread(file = "data/wcvp/wcvp_distribution.csv", header = T)


ng <- read.csv("data/other_countries_floras/sa_2022-07-12_072627325-BRAHMSOnlineData.csv")
nz <- fread("data/other_countries_floras/checklistofthenewzealandflora-seedplantsseptember2020.tsv")
nz2 <- fread("data/other_countries_floras/checklistofthenewzealandflora-fernsandlycophytesseptember2020.tsv")
nc <- read.csv("data/other_countries_floras/Florical_vers_7_I_2022.csv")
indo <- read.csv("data/other_countries_floras/oo_377596.csv")

redlist <- read.csv(file = "data/REDLIST_Folder/simple_summary.csv", header = T)

fivesource <- read.csv(file = "intermediate_data/summary_current.csv", header = T) #Contains outside Aus proportions and endemicity estimates
```

```         
##Adding NZ Data
```

```{r}
#combine nz and nz2 
nz <- rbind(nz, nz2)

#make new df for genus that are possibly endemic (ie score a 1 on genus_end df)
ausplants <- subset(genus_end, prop_endemic == 1)

#remove ? from genus column value in nz data
nz$Genus <- gsub("\\?", "", nz$Genus)

#make a subset for the genus in nz data that appear to have no genus 
no_genera_nz <- subset(nz, Genus == "")
nrow(no_genera_nz) #324 - a problem for another day
nz<-filter(nz, Origin!="Exotic")

ausplants$nz <- ausplants$genus %in% nz$Genus
```

##Adding NC column

```{r}
ausplants$nc <- ausplants$genus %in% nc$Genre
```

##Adding NG column

```{r}
ausplants$ng <- ausplants$genus %in% ng$Genus

```

##Add Indonesian column

```{r}
ausplants$indo <- ausplants$genus %in% indo$Genus
```

###Combining WCVP Distribution and the the Other Thingo

```{r}

wcvp_distribution <- read_csv('data/wcvp/wcvp_distribution.csv')
wcvp <- read_csv('data/wcvp/wcvp_names.csv')

wcvp_comb <- wcvp_distribution %>%
  left_join(wcvp, by = 'plant_name_id') 

wcvp_not_aus <- wcvp_comb %>%
  filter(region != "Australia") #removing all Australian plants

wcvp_aus <- wcvp_comb %>%
  filter(region == "Australia")

sort(unique(wcvp_comb$region)) #does not provide country, but general geographic region
sort(unique(wcvp_aus$area))
print(wcvp_not_aus[wcvp_not_aus$area == "", ])


```

### Adding WCVP Column

```{r}
ausplants$wcvp <- ausplants$genus %in% 
  wcvp_not_aus$genus 
```

### Adding Threatened Status

```{r}
library(dplyr)
library(data.table)
redlist <- redlist %>%
  filter(genusName %in% fivesource$genus) #Filters redlist data to only include genera within fivesource data

unique(redlist$redlistCategory) 
#[1] "Endangered"                 "Critically Endangered"      "Vulnerable"                
#[4] "Least Concern"              "Near Threatened"            "Lower Risk/near threatened"
#[7] "Data Deficient"             "Extinct"

print(redlist[redlist$redlistCategory == "Extinct", ]) #Is Extinct required, should remove? 

#redlist <- redlist %>%
  #filter(redlistCategory != "Extinct")

unique(redlist$genusName) #`182 genera`

counts <- redlist %>% #Creating counts frame for each threatened status
  group_by(genusName) %>%
  count(redlistCategory) %>%
  pivot_wider(names_from = redlistCategory, values_from = n, values_fill = NA) 

# Join the counts with fivesource.trial based on genera2
genus_with_status <- fivesource %>%
  left_join(counts, by = c("genus" = "genusName")) %>% #Lower Risk/ Near Threatened has one value - Maybe combine with Near Threatened?
  mutate_at(c(15:22), ~replace_na(.,0))

#Maybe add proportions??
```

```{r}
genus_with_status$species_number <- as.numeric(genus_with_status$species_number)

options(max.print = 10000)
print(rowSums(genus_with_status[15:22], na.rm = T) > genus_with_status$species_number)

comparison_result <- rowSums(genus_with_status[15:22], na.rm = TRUE) > genus_with_status$species_number


print(genus_with_status[comparison_result, ])
#Arthrotaxis has 2 vulnerable and 1 endangered, with 2 recorded species
#Genoplesium has one recorded spcies, but has critically endangered and one endangered
#Polypogon has one species but 2 least-concern

print(redlist[redlist$genusName == "Athrotaxis", ]) #Athrotaxis laxifolia, Athrotaxis cupressoides, Athrotaxis selaginoides
print(redlist[redlist$genusName == "Genoplesium", ]) #Genoplesium insigne, #Genoplesium baueri 
print(redlist[redlist$genusName == "Polypogon", ]) #Polypogon viridis, Polypogon monspeliensis
```

## Write CSV

```{r}
write_csv(genus_with_status,"intermediate_data/summary_current.csv")
```

## Adding Austraits data - dispersal + growth form traits - for ALL australian genera

```{r}

ausplants <- read_csv('intermediate_data/five_sources_summary.csv')

#create new df with only 'unclear' genera
unclear <- ausplants[ausplants$current_knowledge == 'unclear', ]

#add reason column, to denote reason for confirming endemic/non-endemic status
unclear$reason <- NA

#add endemism status with reason for uncertainty/assessment
unclear$current_knowledge[1] <- 'endemic' 
unclear$reason[1] <- 'ALA taxonomy error'

unclear$current_knowledge[2] <- 'non-endemic' 
unclear$reason[2] <- 'taxonomy error'

unclear$current_knowledge[3] <- 'endemic' 
unclear$reason[3] <- 'POWO data error'

unclear$current_knowledge[4] <- 'endemic' 
unclear$reason[4] <- 'records from 1950s, PNG occurence from NSW Bot Gardens'

unclear$current_knowledge[5] <- 'endemic' 
unclear$reason[5] <- 'POWO data error'

unclear$current_knowledge[6] <- 'endemic'
unclear$reason[6] <- 'Indo data error'

unclear$current_knowledge[7] <- 'endemic'
unclear$reason[7] <- 'POWO data error'

unclear$current_knowledge[8] <- 'non-endemic'
unclear$reason[8] <- 'NC occurence'

unclear$current_knowledge[9] <- 'non-endemic'
unclear$reason[9] <- 'taxonomic error - synonym Machaerina used'

unclear$current_knowledge[10] <- 'non-endemic'
unclear$reason[10] <- 'PNG native'

unclear$current_knowledge[11] <- 'endemic'
unclear$reason[11] <- 'ALA data/shapefile error'

unclear$current_knowledge[12] <- 'endemic'
unclear$reason[12] <- 'ALA data/shapefile error'

unclear$current_knowledge[13] <- 'endemic'
unclear$reason[13] <- 'ornamental plant in UK'

unclear$current_knowledge[14] <- 'endemic'
unclear$reason[14] <- 'taxonomic uncertainty - WCVP says 4 species, zimbabwe flora mentions only 1 accepted species'

unclear$current_knowledge[15] <- 'endemic'
unclear$reason[15] <- 'ng data error'

unclear$current_knowledge[16] <- 'non-endemic'
unclear$reason[16] <- 'taxonomy error, NZ species'

unclear$current_knowledge[17] <- 'endemic'
unclear$reason[17] <- 'introduced NZ and California'

unclear$current_knowledge[18] <- 'endemic'
unclear$reason[18] <- 'introduced India'

unclear$current_knowledge[19] <- 'non-endemic'
unclear$reason[19] <- 'species found in South America'

unclear$current_knowledge[20] <- 'endemic'
unclear$reason[20] <- 'introduced ornamentals in UK'

unclear$current_knowledge[21] <- 'non-endemic'
unclear$reason[21] <- 'species found in Africa, Americas, South Asia etc'

unclear$current_knowledge[22] <- 'endemic'
unclear$reason[22] <- 'taxonomic confusion, tasmania only'

unclear$current_knowledge[23] <- 'non-endemic'
unclear$reason[23] <- 'species found in Africa, Americas, South Asia etc'

unclear$current_knowledge[24] <- 'endemic'
unclear$reason[24] <- 'taxonomic error in indo data'

unclear$current_knowledge[25] <- 'non-endemic'
unclear$reason[25] <- 'found in americas'

unclear$current_knowledge[26] <- 'non-endemic'
unclear$reason[26] <- 'found in PNG POWO taxonomic confusion'

unclear$current_knowledge[27] <- 'endemic'
unclear$reason[27] <- 'POWO error'

unclear$current_knowledge[28] <- 'non-endemic'
unclear$reason[28] <- 'taxonomic uncertainty'

unclear$current_knowledge[29] <- 'non-endemic'
unclear$reason[29] <- 'dist in South America'

unclear$current_knowledge[30] <- 'non-endemic'
unclear$reason[30] <- 'dist in PNG'

unclear$current_knowledge[31] <- 'endemic'
unclear$reason[31] <- 'introdced to india - POWO mismatch'

unclear$current_knowledge[32] <- 'endemic'
unclear$reason[32] <- 'ALA error (Georgiella species entered as Georgeantha'

unclear$current_knowledge[33] <- 'non-endemic'
unclear$reason[33] <- 'Madagascar distribution'

unclear$current_knowledge[34] <- 'endemic'
unclear$reason[34] <- 'introdced to New Zealand - POWO mismatch'

unclear$current_knowledge[35] <- 'non-endemic'
unclear$reason[35] <- 'taxonomic confusion - synonym w justicia'

unclear$current_knowledge[36] <- 'endemic'
unclear$reason[36] <- 'endemic to lord howe - record in NZ?'

unclear$current_knowledge[37] <- 'endemic'
unclear$reason[37] <- 'taxonomic confusion - unplaced in POWO'

unclear$current_knowledge[38] <- 'endemic'
unclear$reason[38] <- 'ALA data error'

unclear$current_knowledge[39] <- 'endemic'
unclear$reason[39] <- 'ALA data error'

unclear$current_knowledge[40] <- 'non-endemic'
unclear$reason[40] <- 'distribution in Chile'

unclear$current_knowledge[41] <- 'endemic'
unclear$reason[41] <- 'ALA data error'

unclear$current_knowledge[42] <- 'non-endemic'
unclear$reason[42] <- 'species found in Africa, Asia, Australia'

unclear$current_knowledge[43] <- 'endemic'
unclear$reason[43] <- 'Indo data error'

unclear$current_knowledge[44] <- 'non-endemic'
unclear$reason[44] <- 'species endemic to Norfolk Island'

unclear$current_knowledge[45] <- 'endemic'
unclear$reason[45] <- 'introduced to NZ and India'

unclear$current_knowledge[46] <- 'endemic'
unclear$reason[46] <- 'ALA data error'

unclear$current_knowledge[47] <- 'endemic'
unclear$reason[47] <- 'ALA data error'

unclear$current_knowledge[48] <- 'endemic'
unclear$reason[48] <- 'introduced to India'

unclear$current_knowledge[49] <- 'endemic'
unclear$reason[49] <- 'introduced to Czechia'

unclear$current_knowledge[50] <- 'non-endemic'
unclear$reason[50] <- 'species found in Americas and Australia'

unclear$current_knowledge[51] <- 'non-endemic'
unclear$reason[51] <- 'species found in Europe, Africa, Asia, Australia'

unclear$current_knowledge[52] <- 'endemic'
unclear$reason[52] <- 'WVCP data error'

unclear$current_knowledge[53] <- 'endemic'
unclear$reason[53] <- 'ALA data error'

unclear$current_knowledge[54] <- 'endemic'
unclear$reason[54] <- 'introduced to Great Britain'

unclear$current_knowledge[55] <- 'endemic'
unclear$reason[55] <- 'introduced to Great Britain'

unclear$current_knowledge[56] <- 'non-endemic'
unclear$reason[56] <- 'synonym for Ochrosia, found across SE & E asia' #Neisosperma

unclear$current_knowledge[57] <- 'non-endemic'
unclear$reason[57] <- 'synonym for Apostasia, found in SE Asia' #Niemeyera

unclear$current_knowledge[58] <- 'endemic'
unclear$reason[58] <- 'Limited PNG occurrence - introduced, flagged as basis of record invalid' #Oraniopsis

unclear$current_knowledge[59] <- 'non-endemic'
unclear$reason[59] <- 'Known to be native to Bolivia and Chile' #Orites

unclear$current_knowledge[60] <- 'non-endemic'
unclear$reason[60] <- 'Known to be native to Mexico and South America' #Orthrosanthus

unclear$current_knowledge[61] <- 'non-endemic'
unclear$reason[61] <- 'Cosmopolitan genera - synonym: Hemionitis' #Paraceterach

unclear$current_knowledge[62] <- 'endemic'
unclear$reason[62] <- 'Data error - single preserved specimen from PNG' #Peripentadenia

unclear$current_knowledge[63] <- 'non-endemic'
unclear$reason[63] <- 'synonym for Arthrochilus- single sp. in PNG' #Phoringopsis

unclear$current_knowledge[64] <- 'non-endemic'
unclear$reason[64] <- 'synonym for Phreatia' #Plexaure

unclear$current_knowledge[65] <- 'non-endemic'
unclear$reason[65] <- 'synonym for Caltha' #Psychrophila

unclear$current_knowledge[66] <- 'endemic'
unclear$reason[66] <- 'WCVP Data error - single 200 year old record in NZ' #Pyrorchis

unclear$current_knowledge[67] <- 'endemic'
unclear$reason[67] <- 'Data duplication of incorrect point' #Rhinerrhiza

unclear$current_knowledge[68] <- 'non-endemic'
unclear$reason[68] <- 'synonym for Phreatia' #Rhipidorchis

unclear$current_knowledge[69] <- 'endemic'
unclear$reason[69] <- 'Introduced and grown as an ornamental' #Rhodanthe

unclear$current_knowledge[70] <- 'endemic'
unclear$reason[70] <- 'Introduced into NZ, UK, & USA' #Rumicastrum

unclear$current_knowledge[71] <- 'non-endemic'
unclear$reason[71] <- 'Synonym with Goodenia'

unclear$current_knowledge[72] <- 'endemic'
unclear$reason[72] <- 'wcvp incorrect, found in PNG'

unclear$current_knowledge[73] <- 'endemic'
unclear$reason[73] <- 'Taxonomy mix up'

unclear$current_knowledge[74] <- 'non-endemic'
unclear$reason[74] <- 'Found in PNG and Malaysia as well'

unclear$current_knowledge[75] <- 'non-endemic'
unclear$reason[75] <- 'Synonym with Brachypterum, found in southern Asia'

unclear$current_knowledge[76] <- 'endemic'
unclear$reason[76] <- 'Outlier in Indo data, only one point'

unclear$current_knowledge[77] <- 'endemic'
unclear$reason[77] <- 'Introduced to Britain'

unclear$current_knowledge[78] <- 'endemic'
unclear$reason[78] <- 'Small numbers outside, botanical gardens'

unclear$current_knowledge[79] <- 'non-endemic'
unclear$reason[79] <- 'Synonym with Aleurites'

unclear$current_knowledge[80] <- 'non-endemic'
unclear$reason[80] <- 'Synonym with Ardisia'

unclear$current_knowledge[81] <- 'non-endemic'
unclear$reason[81] <- 'Found in South America as well'

unclear$current_knowledge[82] <- 'endemic'
unclear$reason[82] <- 'Error in occurrence data, single occurrences outside Australia (oriental/botanical gardens)'

unclear$current_knowledge[83] <- 'endemic'
unclear$reason[83] <- 'taxonomic within genus'

unclear$current_knowledge[84] <- 'non-endemic'
unclear$reason[84] <- 'Native to US and found else where in the world'

unclear$current_knowledge[85] <- 'non-endemic'
unclear$reason[85] <- 'Found in NZ, ALA incorrect'

unclear$current_knowledge[86] <- 'non-endemic'
unclear$reason[86] <- 'Found in New Caledonia and Papua New Guinea'

unclear$current_knowledge[87] <- 'non-endemic'
unclear$reason[87] <- 'Found in Myanmar'

#remove unclears from ausplants df
ausplants <- ausplants[ausplants$current_knowledge != "unclear", ]

#add back now classified 'unclear' genera
ausplants <- rbind(ausplants, unclear[1:13])

#order alphabetically
ausplants <- ausplants[order(ausplants$genus), ]

#rename current_knowlege column to endemism_status
colnames(ausplants)[colnames(ausplants) == 'current_knowledge'] <- 'endemism_status'


write.csv(ausplants, 'intermediate_data/summary_endemism.csv')
```

```{r}
##create df with all australian genera for comparison

library(austraits)

aus_genera <- read_csv('data/australian_flora/genus_level_endemism_estimate.csv')

ausplants <- read_csv('intermediate_data/summary_endemism.csv')

#create non-endemic subset where proportion endemic aus species was <1 (from 2022 Big Data Project)
non_end <- subset(aus_genera, prop_endemic < 1)

#create endemism classification column
non_end$endemism_status <- 'non-endemic'


ausplants <- rbind(non_end[ , c(1,4)], ausplants[ , c(1,13)])

#load austraits
devtools::install_github("traitecoevo/austraits")
austraits <- load_austraits(version = "4.1.0", path = "intro/downloads")

#make genus_list from our australian genera df
genus_list <- ausplants$genus

#now select the categorical traits we want to extract - dispersal appendage, syndrome, plant growth form using the extract_trait function
cat_traits <- austraits %>% extract_trait(c('dispersal_appendage', 'dispersal_syndrome', 'plant_growth_form'))

#join traits to produce dataframe 
cat_traits_joined <- left_join(cat_traits[["traits"]], cat_traits[["taxa"]], by = "taxon_name")

#keep only records for genera on our australian genera list
cat_traits_joined <- cat_traits_joined[cat_traits_joined$genus %in% genus_list, ]

#pivot df such that each row is a genus from the australian genera list, and the most commonly recorded plant growth, dispersal appendage, dispersal syndrome traits are applied for each genus
austraits_genera <- cat_traits_joined %>%
  select(genus, trait_name, value) %>%
  pivot_wider(names_from = trait_name, values_from = value, values_fn = max)

#now the same for seed_traits
disp_traits <- austraits %>% extract_trait(c('seed_dry_mass', 'plant_height'))

#seed traits joined
disp_traits_joined <- left_join(disp_traits[["traits"]], disp_traits[["taxa"]], by = "taxon_name")

#keep only records for genera on our australian genera list
disp_traits_joined <- disp_traits_joined[disp_traits_joined$genus %in% genus_list, ]

#pivot df such that each row is a genus from the australian genera list, and the mean of recorded dry_seed_mass are applied for each genus
disp_genera <- disp_traits_joined %>%
  select(genus, trait_name, value) %>%
  pivot_wider(names_from = trait_name, values_from = value, values_fn = mean)


#add categorical traits and numerical seed mass traits together
austraits_genera <- left_join(austraits_genera, disp_genera, by = 'genus')

write_csv(austraits_genera, 'intermediate_data/all_gen_with_traits.csv')

```

wilcox rank sum tests for plant height and seed mass

```{r}

#contains australian genera and selected traits
austraits_genera <- read_csv('intermediate_data/all_gen_with_traits.csv')

#adding the endemism status to this dataframe for subsequent plotting and tests
genera <- merge(austraits_genera, ausplants[ , c(1,2)], by = 'genus', all.x = TRUE)


library(rstatix)
#wilcox rank sum test for seed_dry_mass
seed_test <- genera %>% 
  wilcox_test(seed_dry_mass ~ endemism_status) %>%
  add_significance()
seed_test #p < 0.0.000733, significant

#wilcox rank sum test for plant_height
height_test <- genera %>% 
  wilcox_test(plant_height ~ endemism_status) %>%
  add_significance()
height_test #p < 0.0.0005, significant

```

Seed mass plots

```{r}
#plotting of seed_dry_mass - violin plot w boxplot
violin_seed <- ggplot(genera, aes(x = endemism_status, y = seed_dry_mass, fill = endemism_status)) +
  geom_violin() +
  scale_fill_brewer() +
  labs(fill = 'Endemic status', x = 'Endemic status', y = 'log dry seed mass (mg)', title = 'Mean seed mass of endemic and non-endemic genera') + 
  theme_bw() +
  #scale_y_log10() +
  scale_y_log10(labels = function(x) sprintf("%.2f", x)) +
  geom_boxplot(width=0.05) 

violin_seed

ggsave('violin_seed.png', violin_seed, width = 7, height = 5)

#plotting of seed_dry_mass - boxplot
ggplot(genera, aes(x = endemism_status, y = seed_dry_mass, fill = endemism_status)) +
  geom_boxplot() +
  scale_fill_brewer() +
  labs(fill = 'Endemic status', x = 'Endemic status', y = 'log dry seed mass (mg)') + scale_y_continuous(labels = scales::number_format(scale = 1e05))+ 
  theme_bw()  


```

Plant height plots

```{r}
#plant height plots - violin w boxplot
violin_height <- ggplot(genera, aes(x = endemism_status, y = plant_height, fill = endemism_status)) +
  geom_violin() +
  scale_fill_brewer() +
  labs(title = 'Mean plant height of endemic and non-endemic genera', fill = 'Endemic status', x = 'Endemic status', y = 'log plant height (cm)') + 
  theme_bw()  +
  scale_y_log10() +
  geom_boxplot(width=0.05) 

violin_height

ggsave('violin_height.png', violin_height, width = 7, height = 5)

#plant height plots - boxplot
ggplot(genera, aes(x = endemism_status, y = plant_height, fill = endemism_status)) +
  geom_boxplot() +
  scale_fill_brewer() +
    scale_y_log10() +
  labs(title = 'Plant height (log) of endemic and non-endemic genera', fill = 'Endemic status', x = 'Endemic status', y = 'plant height (cm)') + 
  theme_bw() 
```

Plotting plant growth form - proportionally

```{r}
#plot growth form - yields 25 different categories - not great visualisation of data
ggplot(genera, aes(x = endemism_status, fill = plant_growth_form)) +
  geom_bar() +
  theme_bw()

#simplifying growth form to combine categories which begin with climber, fern, graminoids, shrub/include shrub, tree, herb. new column growth_form_simp
genera <- genera %>%
  mutate(growth_form_simp = case_when(
    startsWith(plant_growth_form, "climber") ~ "climber",
    startsWith(plant_growth_form, "fern") ~ "fern",
    startsWith(plant_growth_form, "graminoid") ~ "graminoid",
    startsWith(plant_growth_form, "shrub") ~ "shrub",
    grepl("shrub", plant_growth_form, ignore.case = TRUE) ~ "shrub",
    startsWith(plant_growth_form, "tree") ~ "tree",
    startsWith(plant_growth_form, "herb") ~ "herb",
    TRUE ~ plant_growth_form
  ))


#calculating proportion of each simplified growth form
proportions_data <- genera %>%
  group_by(endemism_status, growth_form_simp) %>%
  summarize(count = n()) %>%
  group_by(endemism_status) %>%
  mutate(proportion = count / sum(count))

#remove graminoids, lycophyte, palmoids for plotting as these contain ~1 species 
proportions  <- proportions_data  %>%
  filter(!(growth_form_simp %in% c("palmoid", "lycophyte", "graminoid")))


growth_bar <- ggplot(proportions, aes(x = endemism_status, y = proportion, fill = growth_form_simp)) +
  geom_bar(stat = "identity") +
  labs(x = "Endemism Status", y = "Proportion of genera", title = "Plant growth forms of endemic and non-endemic genera", fill = 'Plant growth form') +
  theme_bw() +
  scale_fill_brewer(palette = "Spectral")

growth_bar 

ggsave('growth_form_bar.png', growth_bar, width = 7, height = 5)
```

Chi square test for plant growth form

```{r}
#contingency table for growth form
cont_form <- table(genera$endemism_status, genera$growth_form_simp)

#chi square test
chi <- chisq.test(cont_form, correct = F)


```

```{r}

#calculating proportion of each simplified growth form
proportions_threat <- all %>%
  group_by(endemism_status, growth_form_simp) %>%
  summarize(count = n()) %>%
  group_by(endemism_status) %>%
  mutate(proportion = count / sum(count))

#remove graminoids, lycophyte, palmoids for plotting as these contain ~1 species 
proportions  <- proportions_data  %>%
  filter(!(growth_form_simp %in% c("palmoid", "lycophyte", "graminoid")))


growth_bar <- ggplot(proportions, aes(x = endemism_status, y = proportion, fill = growth_form_simp)) +
  geom_bar(stat = "identity") +
  labs(x = "Endemism Status", y = "Proportion of genera", title = "Plant growth forms of endemic and non-endemic genera", fill = 'Simplified growth form') +
  theme_bw() +
  scale_fill_brewer(palette = "Set2")

growth_bar 

ggsave('intermediate_data/growth_form_bar.png')
```

```{r}
threat_all <- all %>%
  pivot_longer(cols = c("Least Concern", "Critically Endangered", "Endangered",
                        "Near Threatened", "Vulnerable", "Data Deficient", "Lower Risk"),
               names_to = "threat_status",
               values_to = "count") %>%
  select(endemism_status, threat_status, count)

threat_all <- all %>%
  group_by(endemism_status) %>%
  summarize(across(matches("Least Concern|Critically Endangered|Endangered|Near Threatened|Vulnerable|Data Deficient|Lower Risk"), sum, na.rm = TRUE)) %>%
  pivot_longer(cols = matches("Least Concern|Critically Endangered|Endangered|Near Threatened|Vulnerable|Data Deficient|Lower Risk"),
               names_to = "threat_status",
               values_to = "count")

threat_all <- threat_all %>%
  group_by(endemism_status) %>%
  mutate(proportions = count / sum(count))

threat_order <- c("Critically Endangered", "Endangered", "Vulnerable", "Near Threatened", "Lower Risk", "Least Concern", "Data Deficient")

# Convert threat_status to a factor with the defined order
threat_all$threat_status <- factor(threat_all$threat_status, levels = threat_order)

threat_bar <- ggplot(threat_all, aes(x = endemism_status, y = proportions, fill = threat_status)) +
  geom_bar(stat = "identity") +
  labs(x = "Endemism Status", y = "Proportion of genera", title = "IUCN threatened status of endemic and non-endemic genera", fill = 'IUCN threatened status') +
  theme_bw() +  
  scale_fill_brewer(palette = "RdYlGn")

threat_bar

ggsave('threat_bar.png', threat_bar, width = 7, height = 5)

```
